Design and Implementation of Probabilistic Programming Language Anglican

### 2016-11-27 paper/paper.tex
edits to section on random processes

Similar to distributions, \textit{random processes}, related to
so called `exchangeable random procedures'\footnote{However, random sequences
generated by Anglican random processes are not required to be
exchangeable.}, are defined using
macro \texttt{defproc} and implement the
\texttt{anglican.runtime/{\linebreak[0]}random-process} protocol (that is,
an interface for these random-sequence objects). The protocol
has two methods --- \texttt{produce}, which returns a
distribution object, and \texttt{absorb}, which returns a new
random process object obtained by absorbing a value drawn from
the process. Here is a definition of a beta Bernouilli process:
\begin{lstlisting}[style=default]
(defproc beta-bernoulli
  "beta-Bernoulli process"
  [a b]
  (produce [this] (bernoulli (/ a (+ a b))))
  (absorb [this x]
    (case x
      0 (beta-bernoulli a (inc b))
      1 (beta-bernoulli (inc a) b))))
\end{lstlisting}
While distributions are random and non-functional,
random processes are deterministic and functional. Hence,
\texttt{produce} and \texttt{absorb} are called directly and do
not have corresponding special forms in Anglican:
\begin{itemize}
    \item Values produced by the process are sampled from the distribution
        returned by \texttt{produce}.
    \item When a value is observed from the distribution, a new
        process, with the observed value absorbed, is returned
        by \texttt{absorb}.
\end{itemize}
A common pattern of programming with processes in Anglican is
recursively computing a new process based on the observed value:
\begin{lstlisting}[style=default]
(loop [process (beta-bernoulli 1.0 1.0)
       [value0 & values0] values] 
  (let [dist (produce process)]
    (observe dist value0)
    (recur (absorb process value0) values0)))
\end{lstlisting}
Unlike conventional implementations of random processes,
Anglican's random processes do not have mutable state.

>>>>

In addition to distributions, Anglican provides \textit{random processes},
which define sequences of random variables that are not independent and
identically distributed. Random processes are similar to the so called
`exchangeable random procedures' in Church \cite{GMR+08} and
Venture \cite{MSP14}. However, random sequences generated by Anglican random 
processes are not required to be exchangeable. Random processes are
defined using the \texttt{defproc} macro and implement the
\texttt{anglican.runtime/{\linebreak[0]}random-process} protocol. This
protocol has two methods
\begin{itemize}
\item \texttt{produce}, which returns the distribution on the next
random variable in the sequence, and
\item \texttt{absorb}, which incorporates the value for the next random
variable and returns an updated random process.
\end{itemize}
 As an example, here is a definition of a beta-Bernouilli process, in which
each random variable is distributed according to a Bernoulli distribution with
an unknown parameter that is drawn from a beta distribution:
\begin{lstlisting}[style=default]
(defproc beta-bernoulli
  "beta-Bernoulli process"
  [a b]
  (produce [this] (bernoulli (/ a (+ a b))))
  (absorb [this x]
    (case x
      0 (beta-bernoulli a (inc b))
      1 (beta-bernoulli (inc a) b))))
\end{lstlisting}
Unlike typical implementations of exchangeable random processes, Anglican's random processes do not have mutable state. The \texttt{produce} and \texttt{absorb} methods are deterministic and functional, and therefore do not have corresponding special forms in Anglican. Random values are generated using a recursion in which \texttt{produce} constructs the distribution on the next variable and \texttt{absorb} constructs the updated process for the next iteration. For example:
\begin{lstlisting}[style=default]
(defm sample-beta-binomial [n a b]
  (loop [process (beta-bernoulli a bb)
         values []]
    (if (= (count values) n)
      values
      (let [dist (produce process)
            value (sample dist)]
      (recur (absorb process value)
             (conj values value)))))))
\end{lstlisting}
Similarly, random processes can also be used to recursively observe a sequence of values:
\begin{lstlisting}[style=default]
(loop [process (beta-bernoulli 1.0 1.0)
       [value0 & values0] values] 
  (let [dist (produce process)]
    (observe dist value0)
    (recur (absorb process value0) values0)))
\end{lstlisting}

% \begin{itemize}
%     \item Values produced by the process are sampled from the distribution
%         returned by \texttt{produce}.
%     \item When a value is observed from the distribution, a new
%         process, with the observed value absorbed, is returned
%         by \texttt{absorb}.
% \end{itemize}

% A common pattern when programming with random processes in 
% Anglican is to generate a sequences of sampled values and 
% constructing an updated process at each step
